---
title: "Quick Start: Applying the Cell Key Method"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{quickstart-ckm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ckm)
```

The package provides a sample dataset for practice:
```{r data}
data("dtest", package = "ckm")
```

## Step 1: Add Individual Keys

The function `build_individual_keys()` adds a column `rkey` to your dataset, containing individual keys drawn from a uniform distribution between 0 and 1.

```{r cles}
set.seed(451919)
dtest_with_keys <- build_individual_keys(dtest)
```

You can check that the distribution of individual keys is uniform:

```{r unif}
hist(dtest_with_keys$rkey)
```

## Apply the Cell Key Method in One Step

After generating the individual key, you can directly build the perturbed table.

The code below will:

- Build an aggregated table crossing the variables `"DEP", "DIPLOME", "SEXE", "AGE"`, including margins.
- Compute the perturbation table according to parameters `D` (maximum deviation) and `V` (variance of the distribution).
- Apply the perturbation to the aggregated table.

```{r ckm_direct}
res_ckm <- tabulate_and_apply_ckm(
 df = dtest_with_keys,
 cat_vars = c("DEP", "DIPLOME", "SEXE", "AGE"),
 D = 5, V = 2
)
```

The result is a list with four elements. The first element is the perturbed table, the second is the transition matrix, the third is the risk and utility measures, and the fourth is the original table.
You can check the structure of the perturbed table:

```{r structure-res-ckm} 
str(res_ckm$tab)
```

## Apply the Cell Key Method Step by Step

### Step 1 Generate the Counting Table with Cell Keys

The next step is to build your counting table. You need to provide the dataset with the individual keys and the categorical variables you want to cross.

The function `tabulate_cnt_micro_data` builds all possible cross-tabulations (including margins) and provides both the counts and the cell key (the decimal part of the sum of individual keys).

```{r tabul}
tab_before <- tabulate_cnt_micro_data(
 df = dtest_with_keys,
 cat_vars = c("DEP", "DIPLOME", "SEXE", "AGE")
)
```

The `marge_label` argument lets you set the margin label (default is `"Total"`), and the `hrc_vars` argument allows you to specify a hierarchy between variables if needed.

### Step 2 Apply the Perturbation

The next step is to apply the perturbation to the previous table. You need to specify at least:

- The table to perturb
- The two main CKM parameters:
  - Maximum deviation `D` (e.g., `D=5` means data are perturbed by an integer between `-5` and `+5`)
  - Variance `V` of the probability distribution. The higher the variance, the greater the overall perturbation (e.g., measured by the mean absolute deviation).

```{r appliquer}
res_ckm <- apply_ckm(tab_before, D = 5, V = 2)
```

```{r structure-res}
str(res_ckm$tab)
```
