---
title: "Applying the Cell Key Method with the `ckm` package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using-ckm-package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates how to use the key functions of the package to apply the Cell Key Method for statistical disclosure control. The steps include:

1. Adding keys to microdata using `build_individual_keys`.
2. Comparing different transition matrices with `visualiser_distribution` for various parameter choices.
3. Simulating risk and utility measures with `simulate_RUs`.
4. Applying the Cell Key Method with `tabulate_and_apply_ckm` and interpreting the results.

# Step 1: Adding Keys to Microdata

First, load the package and example data, then generate individual keys.

```{r setup-load-data}
library(dplyr)
library(ckm)

# Load example microdata
data("dtest", package = "ckm")

# Set a seed for reproducibility
set.seed(4081789)

# Add individual keys to the microdata
dtest_with_keys <- build_individual_keys(dtest)
head(dtest_with_keys)
```

# Step 2: Comparing Transition Matrices

## Step 2.1: Search Feasible Transition Matrices

```{r}
test_matrices(D = 10, js = 4)
```

```{r}
test_matrices(D = 15, js = 4)
```

With D equals to 10 or 15 and js equals to 4, the minimal variance that let us 
build a transition matrix is`V=6.5625`.


## Step 2.2: Visualize the distributions

We compare the transition matrices for different values of D and V using `visualiser_distribution`.

```{r compare-transition-matrices}
# Visualize transition matrices for different parameter sets
visualize_distribution(D = c(10, 15), V = c(10, 20, 30))
```

With `V<=20`, the distributions ensures that more than `75%` of the absolute 
deviations are equal to 5 at most.

# Step 3: Estimating Risk and Simulating Utility Measures

We use `simulate_RUs` to compare the risk and utility for the different parameter sets. 
We set `n_sim = 5` for a quick demonstration and use the categorical variables REG, DIPLOME, AGE, and SEXE.

```{r simulate-risk-utility}
# Define the categorical variables
cat_vars <- c("REG", "DIPLOME", "AGE", "SEXE")
# Define the parameters D and V (and js if needed) to test:
parameters <- build_parameters_table(Ds = c(10,15), Vs = c(10, 20), jss = 4)

# Simulate risk and utility for different parameter sets
sim_results <- simulate_RUs(
  df = dtest,
  cat_vars = cat_vars,
  parametres = parameters,
  confident = 5,
  n_sim = 5
)

# Display the results
sim_results |> 
  group_by(D, V, js, i, j, risk_inference = qij) |>
  summarise(MAD = mean(MAD), .groups = "drop")
```

As the risk inference of sensitive values is lower than `50%` for all sets of 
parameters, one can choose the one that maximize the utility (`D=15, V=10`).

# Step 4: Applying the Cell Key Method

Based on the previous results, we select one set of parameters 
(D = 15, V = 10) and apply the Cell Key Method using 
`tabulate_and_apply_ckm`.

```{r apply-ckm}
# Apply the Cell Key Method with chosen parameters
res_ckm <- tabulate_and_apply_ckm(
  df = dtest_with_keys,
  cat_vars = cat_vars,
  D = 15,
  V = 10,
  js = 4
)

# Examine the perturbed table
head(res_ckm$tab |> filter(DIPLOME == "D") |> arrange(REG, AGE, SEXE), n = 12)
```

